Intelligent Building Energy Management System

### 2021-05-07 parts/10-introduction.tex
Updated report

During a research project led by Miah \textit{et al.,} the DC motor interface was successfully
implemented into the BEMOSS platform which required adding a great deal of source code
to the platform to fully implement the device as BEMOSS does not support adding
new devices currently. The research also consisted of adding a new
features such a logo to the
platform for the motor. A possible improvement our platform will make is helping
developers on the project to add new devices more easily.

@@

One such platform is the BEMOSS from the Virginia Polytechnic Institute and State
University Advanced Research Institute. BEMOSS features include being
open-source, allowing for multiple communication technologies, and supporting
many common IoT devices. This system allows for monitoring and controlling a
variety of devices securely~\cite{BEMOSS}.

Another open-source platform is rEMpy or residential Energy Management in
Python. The main focus of this platform developed by a team of researchers at
Universit\`{a} Politecnica delle Marche in Ancona, Italy was to simulate the
energy flow of a residential home. With the data collected from simulations, it
was possible for them to use forecasting and prediction algorithms to predict
various quantities such as energy usage and power consumption over a given time period. Their energy management system consists of a well-defined structure with components like a user interface, database, and optimal scheduler that communicates with modules like a prediction and thermal model. In terms of software technologies, it uses the Django web framework which provides a lot of built-in functionality for web development like a user-authentication system and object-relational mapping. Adapting the project to this framework could be a great way to help improve security and optimize database queries~\cite{fagiani2017}.

@@

In~\cite{8246800}, authors describe similar functional requirements to our
own BEMS Core. They periodically gather power consumption and device status data
and send that data to a database. We are doing exactly what they describe using
the Apache Cassandra database. The authors also use an \say{analytics [sic] engine
to process [the data] and generate reports, graphs, and charts.} Our system only
generates a power consumption plot for a given day, but we are still meeting
this same requirement - just on a smaller scale. The communication with devices
and power consumption reports are to be accessible through a web-based
application that is easy to use. Again, our implementation is simple and very
intuitive for the end-user. Finally, these authors mention a user-login feature.
They explain that their system would include multiple levels of accessibility
based on the user privileges. Although our system does not currently have a
user-login feature, it simply uses a router to access devices that are also
connected to the router. Since this is a closed system, users need only protect
their router with a secure password.

>>>>

During a research project led by Miah \textit{et al.,} the DC motor interface was successfully
implemented into the BEMOSS platform which required adding a great deal of source code
to the platform to fully implement the device as BEMOSS does not support adding
new devices currently. The research also consisted of adding new
features such as a logo to the
platform for the motor. A possible improvement our platform will make is helping
developers on the project add new devices more easily.

@@

One such platform is BEMOSS from the Virginia Polytechnic Institute and State
University Advanced Research Institute. BEMOSS features include being
open-source, allowing for multiple communication technologies, and supporting
many common IoT devices. This system allows for monitoring and controlling a
variety of devices securely~\cite{BEMOSS}.

Another open-source platform is rEMpy or residential Energy Management in
Python. The main focus of this platform developed by a team of researchers at
Universit\`{a} Politecnica delle Marche in Ancona, Italy was to simulate the
energy flow of a residential home. With the data collected from simulations, it
was possible for them to use forecasting and prediction algorithms to predict
various quantities such as energy usage and power consumption over a given time period. Their energy management system consists of a well-defined structure with components like a user interface, database, and optimal scheduler that communicate with modules like a prediction and thermal model. In terms of software technologies, it uses the Django web framework which provides a lot of built-in functionality for web development like a user-authentication system and object-relational mapping. Adapting the project to this framework could be a great way to help improve security and optimize database queries~\cite{fagiani2017}.

@@

In~\cite{8246800}, authors describe similar functional requirements to our
own BEMS Core. They periodically gather power consumption and device status data
and send that data to a database. We are doing exactly what they describe using
the Apache Cassandra database. The authors also use an \say{analytics [sic] engine
to process [the data] and generate reports, graphs, and charts.} Our system only
generates a power consumption plot for a given day, but we are still meeting
this same requirement - just on a smaller scale. The communication with devices
and power consumption reports are to be accessible through a web-based
application that is easy to use. Again, our implementation is simple and very
intuitive for the end-user. Finally, these authors mention a user-login feature.
They explain that their system would include multiple levels of accessibility
based on the user privileges.

### 2021-05-07 parts/20-iBEMSArchitecture.tex
Updated report

The Flask web server is built using a Python web framework titled Flask which
allows programmers to develop a dynamic web server capable of rendering data to
HTML pages. The benefit of this framework over other Python web frameworks like
Django is the large amount of customization available. For example, the
developers can create a completely custom user login system. For now, the server
is capable of being accessed only on the server computer running Ubuntu Linux.
In the future, support could be added to allow any user on the LAN to access the
web server. Other web frameworks use object relational mapping directly to
access the relevant databases. However, our system uses direct queries with both
SQLite and the Cassandra query language to query the database. \medbreak We used
2 different databases to store information on. The Apache Cassandra database was
better suited for the time series data such as device status and power usage.
This database was therefore utilized a lot by the Control and Scheduling Agents.
The SQLite database hold parameters that do not change over time like device
ID's, IP addresses, and MAC addresses which made it amendable to the processes
in the Discovery Agent. A possible improvement to this project could be to only
utilize 1 database for development simplicity. \medbreak The first agent that is
used by the iBEMS Core on startup is the discovery agent. It is responsible for
reaching out on the network and finding all devices that the iBEMS Core has
support for. This process involves retrieving the following parameters: IP
address, Port number, MAC address, Manufacturer, Name, and API. Once all
available devices are connected, the Control agent is used to change the status
of each device and collect power usage data in real time. Finally, the
scheduling will take input from the user and place the corresponding scheduling
periods in the Apache Cassandra database. Then, a thread is created to poll the
current device status at regular intervals and updates the status if it does not
match the defined status in the Apache Cassandra database.

@@

There are 5 general modes of operation in iBEMS as shown in Figure~\ref{fig:operationalModes}. In Mode 0, before iBEMS is even launched, the user must open the Startup Menu, which is a desktop graphical user interface (GUI). This is a simple GUI with only 2 buttons for starting and stopping iBEMS. Upon clicking \say{Start iBEMS}, the login page will load where the user can enter a valid username and password to use iBEMS. Once a valid username and password is entered, the home screen will be shown which displays the weather data being downloaded from the OpenWeatherMap API. This information include temperature in Fahrenheit and Celsius, wind speed in MPH, humidity, and cloud coverage. At this point, a navigation bar is available to view \say{Active Devices} page and the \say{Scheduling} page. \say{Active Devices} corresponds to Mode 3 for controlling devices in real time. The user will also be able to view power plots in this mode as well. Lastly, the \say{Scheduling} page is Mode 4 for creating schedules for all connected devices.

@@

A large feature of this project was being able to record power usage from the
connected IoT devices. The embedded computer does not have a built-in capability
to measure its own power usage, so this circuit was needed to make that
possible. Essentially, this circuit will measure the current being used to drive
one of the motors on the robot chassis. The on-board Analog-to-Digital-Converter
(ADC) can be used to read the voltage across the resistor and since the
resistance is known, the current can be calculated as $I = V/R$. Also, the
voltage from the H-Bridge to drive the motors on the embedded computer is known
so the power consumed by 1 motor is $Motor Drive Voltage * I$. For the specific
robot chassis used in this project, there are 4 motors, so multiplying the power
found from this circuit by 4 will give an approximation of the total power used
by the robot. For other robots with different power sources besides a Lipo
battery. See Figure~\ref{fig:motorInterfaceCircuit} for details of the interface
circuit for computer power usage by the DC motor.

@@

The other supported device is the WeMo Insight Switch as seen in Figure~\ref{fig:wemo}. This device was easy to develop support for as it is meant for building energy management. It simply plugs into a 120 volt 3-prong outlet and any typical household appliance running on 120 volts can be plugged into it. Internal circuitry allows the WeMo to be controlled remotely (turn On/Off) and to record power usage of whatever is plugged into.

>>>>

The Flask web server is built using a Python web framework titled Flask which
allows programmers to develop a dynamic web server capable of rendering data to
HTML pages. The benefit of this framework over other Python web frameworks like
Django is the large amount of customization available. For example, the
developers can create a completely custom user login system. For now, the server
is capable of being accessed only on the server computer running Ubuntu Linux.
In the future, support could be added to allow any user on the LAN to access the
web server. Other web frameworks use object relational mapping directly to
access the relevant databases. However, our system uses direct queries with both
SQLite and the Cassandra query language to query the databases. \medbreak We used
2 different databases to store information on. The Apache Cassandra database was
better suited for the time series data such as device status and power usage.
This database was therefore utilized a lot by the Control and Scheduling Agents.
The SQLite database hold parameters that do not change over time like device
ID's, IP addresses, and MAC addresses which made it amendable to the processes
in the Discovery Agent. A possible improvement to this project could be to only
utilize 1 database for development simplicity. \medbreak The first agent that is
used by the iBEMS Core on startup is the discovery agent. It is responsible for
reaching out on the network and finding all devices that the iBEMS Core has
support for. This process involves retrieving the following parameters: IP
address, Port number, MAC address, Manufacturer, Name, and API. Once all
available devices are connected, the Control agent is used to change the status
of each device and collect power usage data in real time. Finally, the
scheduling will take input from the user and place the corresponding scheduling
periods in the Apache Cassandra database. Then, a thread is created to poll the
current device status at regular intervals and update the status if it does not
match the defined status in the Apache Cassandra database.

@@

There are 5 general modes of operation in iBEMS as shown in Figure~\ref{fig:operationalModes}. In Mode 0, before iBEMS is even launched, the user must open the Startup Menu, which is a desktop graphical user interface (GUI). This is a simple GUI with only 2 buttons for starting and stopping iBEMS. Upon clicking \say{Start iBEMS}, the login page will load where the user can enter a valid username and password to use iBEMS. Once a valid username and password is entered, the home screen will be shown which displays the weather data being downloaded from the OpenWeatherMap API. This information includes temperature in Fahrenheit and Celsius, wind speed in MPH, humidity, and cloud coverage. At this point, a navigation bar is available to view the \say{Active Devices} page and the \say{Scheduling} page. \say{Active Devices} corresponds to Mode 3 for controlling devices in real time. The user will also be able to view power plots in this mode as well. Lastly, the \say{Scheduling} page is Mode 4 for creating schedules for all connected devices.

@@

A large feature of this project was being able to record power usage from the
connected IoT devices. The embedded computer does not have a built-in capability
to measure its own power usage, so a circuit was needed to make that
possible. Essentially, this circuit will measure the current being used to drive
one of the motors on the robot chassis. The on-board Analog-to-Digital-Converter
(ADC) can be used to read the voltage across the resistor and since the
resistance is known, the current can be calculated as $I = V/R$. Also, the
voltage from the H-Bridge to drive the motors on the embedded computer is known
so the power consumed by 1 motor is $Motor Drive Voltage * I$. For the specific
robot chassis used in this project, there are 4 motors, so multiplying the power
found from this circuit by 4 will give an approximation of the total power used
by the robot. \add{For other robots with different power sources besides a Lipo
battery.} See Figure~\ref{fig:motorInterfaceCircuit} for details of the interface
circuit for computer power usage by the DC motor.

@@

The other supported device is the WeMo Insight Switch as seen in Figure~\ref{fig:wemo}. This device was easy to develop support for as it is meant for building energy management. It simply plugs into a 120 volt 3-prong outlet and any typical household appliance running on 120 volts can be plugged into it. Internal circuitry allows the WeMo to be controlled remotely (turn On/Off) and to record power usage of whatever it is plugged into.

### 2021-05-07 parts/30-Implementation.tex
Updated report

\subsection{Discovery Agent Functionality}
The purpose of the discovery agent is to search for devices that respond to the supported APIs by the software. A flow chart demonstrating the \texttt{searchForDevices} method is shown in Figure~\ref{fig:searchForDevices}. This method is called whenever the active devices page is loaded. The algorithm could be improved to help page load times. All the device APIs are loaded in the \texttt{findDevices} function defined in each device's API module is called which will return a list of found URLs. Once the device's URL is determined, it's metadata can be passed to the \texttt{setDeviceToActive} method which will set up the device in the two databases.

\begin{figure}[H]
    \centering
    \includegraphics[scale=0.5]{figs/searchForDevices.pdf}
    \caption{searchForDevices Flow Chart}
    \label{fig:searchForDevices}
\end{figure}
The \texttt{setDeviceToActive} shown in Figure~\ref{fig:setDeviceToActive} will first check to make sure the device is not already in the active devices table. Then, it will add the device to the active devices table, create a table in the Cassandra database for storing time-series data, send a message to the scheduling agent to check for schedule updates, and send a message to the control agent to start collecting data.
\begin{figure}[H]
    \centering
    \includegraphics[scale=0.5]{figs/setDeviceToActive.pdf}
    \caption{setDeviceToActive Flow Chart}
    \label{fig:setDeviceToActive}
\end{figure}

\subsection{Scheduling Agent Functionality}
The scheduling agent is able to update the device status of each device dependent on whether a schedule is set for the specific time period and day. One of the methods that is called, \texttt{updateScheduleFromServer} is invoked when the user presses the \say{Update Schedule} button from the UI on the scheduling page. A flow chart of this method is shown in Figure~\ref{fig:updateScheduleFromServer}. When this method is called the desired schedule is loaded in as a dictionary and parsed to extract the necessary periods and given day. First the times are stored in military time in the database for more convenient storage and parsing later on. Then, a table is created in the Apache Cassandra database for the device for the provided data and the device ID is embedded. The table's data is destroyed to ensure there are no errors regarding overlapping schedules. Lastly, all the scheduling periods extracted from the dictionary are stored as rows in the table. 
\begin{figure}[H]
    \centering
    \includegraphics[scale=0.5]{figs/updateScheduleFromServer.pdf}
    \caption{updateScheduleFromServer Flow Chart}
    \label{fig:updateScheduleFromServer}
\end{figure}
A second method that is run inside each device thread is \texttt{periodicDeviceUpdateMethod}, defined in Figure~\ref{fig:periodicDeviceUpdateMethod}, is started when the discovery agent sets the discovered device to active. By first loading the device's API name, the scheduling agent can begin checking for scheduling periods inside an infinite loop. The status dictionary is queried from the Cassandra database and loaded into a dictionary. All these scheduling periods are looped over for the current day while constantly checking whether the current time in seconds is within the starting and ending time. A conversion of both starting and ending time to seconds is necessary as it is challenging to work with these quantities in standard and military time when comparing times. If the current time does lie inside the scheduling period, the device's state is set dependent on the API. 
\begin{figure}[H]
    \centering
    \includegraphics[scale=0.5]{figs/periodicDeviceUpdateMethod.pdf}
    \caption{periodicDeviceUpdateMethod Flow Chart}
    \label{fig:periodicDeviceUpdateMethod}
\end{figure}

@@

The embedded computer needed some software to be able to connect to and communicate with the iBEMS Core. To achieve this, a program consisting of 2 functions for connecting and communicating is used.
\medbreak
The first of these functions is \texttt{connect\_to\_iBEMS}. This function first turns on the red LED on the embedded computer to indicate it is not yet connected. Then it creates a socket using a multicast IP address that the corresponding API in the iBEMS core will be looking for. Once the socket is created, the embedded computer will enter a loop and wait for the iBEMS core to make the initial connection. When that connection is made, it will exit the loop, save the IP address of the iBEMS core, send a confirmation message, then turn on the green LED before closing the multicast socket.
\medbreak
The second function, \texttt{communicate\_with\_iBEMS} will open another socket but with the IP address of the iBEMS core found in \texttt{connect\_to\_iBEMS}. It will then enter a indefinite loop to continually receive data from the iBEMS core. There are 4 different messages that the iBEMS core sends to the embedded computer. First, it will check if the iBEMS core is telling it to shut down with the command \say{Shutdown}. If so, the embedded computer will immediately shut down. Second, if the data sent by iBEMS is \say{status}, the embedded computer will take a reading from the circuit described in Chapter~\ref{ch: Chapter2} section 3. In case the user decides to shut down the iBEMS core without turning off the embedded computer, another message \say{disconnect} will be sent in which case the embedded computer will exit \texttt{communicate\_with\_iBEMS} and start the entire program over again. Finally, if iBEMS is not sending any of the above options, it means that the user is sending a new duty cycle to drive the motors with. Since iBEMS only sends messages with the string data type, this has to be converted to a float in order to call the motor driver function.

>>>>

\subsection{Discovery Agent Functionality}
The purpose of the discovery agent is to search for devices that respond to the supported APIs by the software. A flow chart demonstrating the \texttt{searchForDevices} method is shown in Figure~\ref{fig:searchForDevices}. This method is called whenever the active devices page is loaded. The algorithm could be improved to help page load times. All the device APIs are loaded by the \texttt{findDevices} function which will return a list of found URLs. Once the device's URL is determined, it's metadata can be passed to the \texttt{setDeviceToActive} method which will set up the device in the two databases.

\begin{figure}[H]
    \centering
    \includegraphics[scale=0.5]{figs/searchForDevices.pdf}
    \caption{searchForDevices Flow Chart}
    \label{fig:searchForDevices}
\end{figure}
The \texttt{setDeviceToActive} method shown in Figure~\ref{fig:setDeviceToActive} will first check to make sure the device is not already in the active devices table. Then, it will add the device to the active devices table, create a table in the Cassandra database for storing time-series data, send a message to the scheduling agent to check for schedule updates, and send a message to the control agent to start collecting data.
\begin{figure}[H]
    \centering
    \includegraphics[scale=0.5]{figs/setDeviceToActive.pdf}
    \caption{setDeviceToActive Flow Chart}
    \label{fig:setDeviceToActive}
\end{figure}

\subsection{Scheduling Agent Functionality}
The scheduling agent is able to update the device status of each device dependent on whether a schedule is set for the specific time period and day. One of the methods that is called, \texttt{updateScheduleFromServer} is invoked when the user presses the \say{Update Schedule} button from the UI on the scheduling page. A flow chart of this method is shown in Figure~\ref{fig:updateScheduleFromServer}. When this method is called the desired schedule is loaded in as a dictionary and parsed to extract the necessary periods and given day. First the times are stored in military time in the database for more convenient storage and parsing later on. Then, a table is created in the Apache Cassandra database where the device ID and day are embedded in the name. The previous table's data is destroyed to ensure there are no errors regarding overlapping schedules. Lastly, all the scheduling periods extracted from the dictionary are stored as rows in the table. 
\begin{figure}[H]
    \centering
    \includegraphics[scale=0.5]{figs/updateScheduleFromServer.pdf}
    \caption{updateScheduleFromServer Flow Chart}
    \label{fig:updateScheduleFromServer}
\end{figure}
A second method that is run inside each device thread is \texttt{periodicDeviceUpdateMethod}, defined in Figure~\ref{fig:periodicDeviceUpdateMethod}, which is started when the discovery agent sets the discovered device to active. By first loading the device's API name, the scheduling agent can begin checking for scheduling periods inside an infinite loop. The status dictionary is queried from the Cassandra database and loaded into a dictionary. All these scheduling periods are looped over for the current day while constantly checking whether the current time in seconds is within the starting and ending time. A conversion of both starting and ending time to seconds is necessary as it is challenging to work with these quantities in standard and military time when comparing times. If the current time does lie inside the scheduling period, the device's state is set dependent on the API. 
\begin{figure}[H]
    \centering
    \includegraphics[scale=0.5]{figs/periodicDeviceUpdateMethod.pdf}
    \caption{periodicDeviceUpdateMethod Flow Chart}
    \label{fig:periodicDeviceUpdateMethod}
\end{figure}

@@

The embedded computer needed some software to be able to connect to and communicate with the iBEMS Core. To achieve this, a program consisting of 2 functions for connecting and communicating is used.
\medbreak
The first of these functions is \texttt{connect\_to\_iBEMS}. This function first turns on the red LED on the embedded computer to indicate it is not yet connected. Then it creates a socket using a multicast IP address that the corresponding API in the iBEMS core will be looking for. Once the socket is created, the embedded computer will enter a loop and wait for the iBEMS core to make the initial connection. When that connection is made, it will exit the loop, save the IP address of the iBEMS core, send a confirmation message, then turn on the green LED before closing the multicast socket.
\medbreak
The second function, \texttt{communicate\_with\_iBEMS} will open another socket but with the IP address of the iBEMS core found in \texttt{connect\_to\_iBEMS}. It will then enter an indefinite loop to continually receive data from the iBEMS core. There are 4 different messages that the iBEMS core sends to the embedded computer. First, it will check if the iBEMS core is telling it to shut down with the command \say{Shutdown}. If so, the embedded computer will immediately shut down. Second, if the data sent by iBEMS is \say{status}, the embedded computer will take a reading from the circuit described in Chapter~\ref{ch: Chapter2} section 3. In case the user decides to shut down the iBEMS core without turning off the embedded computer, another message \say{disconnect} will be sent in which case the embedded computer will exit \texttt{communicate\_with\_iBEMS} and start the entire program over again. Finally, if iBEMS is not sending any of the above options, it means that the user is sending a new duty cycle to drive the motors with. Since iBEMS only sends messages with the string data type, this has to be converted to a float in order to call the motor driver function.
